### 2014-11-05 contract-language.tex
Committing first experiments
Major addition to implementation
Edits throughout

\section{Contract Language}
\label{sec:contract-lang}

@@

\subsection{Capturing Store Semantics}

@@

\item \textbf{Causal consistency}: Causally consistent operations are required
	to see a causally consistent snapshot of the object state, including the
	actions performed on the same session.  The latter requirement implies that
	if two operations $o_1$ and $o_2$ from the same session are applied to two
	different replicas $r_1$ and $r_2$, the second operation cannot be discharged
	until the effect of $o_1$ is merged with $o_2$ in both $r_1$ and $r_2$. The
	\cf{getBalance} operation requires causal consistency, as it requires the
	operations from the same session to be visible, which cannot be guaranteed
	under just eventual consistency. We assume that causality is only tracked
	through operations on the same object. The corresponding store semantics is
	captured by the contract $\ccc$ defined below:

  \vspace{-1em}
  \begin{smathpar}
  \ccc = \forall a.~(\hboZ \cap \sameobjZ) (a,\cureff) \Rightarrow \vis{a}{\cureff}
  \end{smathpar}

@@

% SJ: not sure this is necessary, or appropriate here.
%% \noindent Observe that out contract language does not incorporate real (i.e.,
%% wall-clock) time. Hence, it cannot describe store
%% semantics such as recency or bounded-staleness
%% guarantees offered by certain stores~\cite{Pileus}.

@@

If the contract ($\cv_{op}$) of an operation ($\mathit{op}$) is \emph{weaker
than} a store contract ($\cv_{st}$), then constraints expressed by the
former are implied by guarantees provided by the latter.  The
completeness of first-order logic\cite{completeness} allows us to assert
that any well-formed execution ($\E$) that satisfies $\cv_{st}$ (i.e.,
$\E\models \cv_{st}$) also satisfies $\cv_{op}$ (i.e., $\E \models
\cv_{op}$). Consequently, it is safe to execute operation $\mathit{op}$ under
a store consistency level captured by $\cv_{st}$.

Observe that the contracts $\scc$, $\ccc$ and $\ecc$ are themselves
totally ordered with respect to the $\le$ relation: $\ecc \le \ccc \le
\scc$.  This concurs with the intuition that any contract satisfiable
under $\ecc$ or $\ccc$ is satisfiable under $\scc$, and any contract
that is satisfiable under $\ecc$ is satisfiable under $\ccc$. However,
we are interested in the \emph{weakest} guarantee (among $\ecc$,
$\ccc$, and $\scc$) required to satisfy the contract. We define the
corresponding consistency level as the \emph{consistency class of the
contract}. The classification scheme, presented formally in
Figure~\ref{sem:classify}, defines rules to judge the consistency class of
a contact. For example, the scheme classifies the \cf{getBalance} contract
($\cv_{gb}$) from Section~\ref{sec:motivation} as a {\sf
CausallyConsistent} contract, because the sequent $\Delta \vdash \ccc
\Rightarrow \cv_{gb}$ is valid in first-order logic (therefore,
$\cv_{gb} \le \ccc$), whereas the sequent $\Delta \vdash \ecc
\Rightarrow \cv_{gb}$ is invalid (therefore, $\cv_{gb} \not\le \ecc$).
Due to the confinement of our contract language to a decidable subset
of the logic, validity of such sequents can be decided mechanically
allowing us to automate the classification scheme in \name.

@@

We now present a meta-theoretic result that certifies the soundness of
classification-based contract enforcement. To help us state the result, we
define an axiomatic model of the our system described informally in
Section~\ref{sec:sysmod}:

>>>>

\section{Contract Language}
\label{sec:lang}

@@

\subsection{Capturing Store Semantics}
\label{sec:store_sem}

@@

\item \textbf{Causal consistency}: Causally consistent operations are required
	to see a causally consistent snapshot of the object state, including the
	actions performed on the same session.  The latter requirement implies that
	if two operations $o_1$ and $o_2$ from the same session are applied to two
	different replicas $r_1$ and $r_2$, the second operation cannot be discharged
	until the effect of $o_1$ is merged with $o_2$ in both $r_1$ and $r_2$. The
	\cf{getBalance} operation requires causal consistency, as it requires the
	operations from the same session to be visible, which cannot be guaranteed
	under just eventual consistency. The corresponding store semantics is
	captured by the contract $\ccc$ defined below:

  \vspace{-1em}
  \begin{smathpar}
  \ccc = \forall a.~\hbo{a}{\cureff} \Rightarrow \vis{a}{\cureff}
  \end{smathpar}

@@

@@

If the contract ($\cv_{op}$) of an operation ($\mathit{op}$) is \emph{weaker
than} a store contract ($\cv_{st}$), then constraints expressed by the former
are implied by guarantees provided by the latter. The completeness of
first-order logic allows us to assert that any well-formed execution ($\E$)
that satisfies $\cv_{st}$ (i.e., $\E\models \cv_{st}$) also satisfies
$\cv_{op}$ (i.e., $\E \models \cv_{op}$). Consequently, it is safe to execute
operation $\mathit{op}$ under a store consistency level captured by $\cv_{st}$.

Observe that the contracts $\scc$, $\ccc$ and $\ecc$ are themselves totally
ordered with respect to the $\le$ relation: $\ecc \le \ccc \le \scc$.  This
concurs with the intuition that any contract satisfiable under $\ecc$ or $\ccc$
is satisfiable under $\scc$, and any contract that is satisfiable under $\ecc$
is satisfiable under $\ccc$. However, we are interested in the \emph{weakest}
guarantee (among $\ecc$, $\ccc$, and $\scc$) required to satisfy the contract.
We define the corresponding consistency level as the \emph{consistency class of
the contract}. The classification scheme, presented formally in
Figure~\ref{sem:classify}, defines rules to judge the consistency class of a
contact. For example, the scheme classifies the \cf{getBalance} contract
($\cv_{gb}$) from \S~\ref{sec:motivation} as a {\sf CausallyConsistent}
contract, because the sequent $\Delta \vdash \ccc \Rightarrow \cv_{gb}$ is
valid in first-order logic (therefore, $\cv_{gb} \le \ccc$), whereas the
sequent $\Delta \vdash \ecc \Rightarrow \cv_{gb}$ is invalid (therefore,
$\cv_{gb} \not\le \ecc$). Due to the confinement of our contract language to a
decidable subset of the logic, validity of such sequents can be decided
mechanically allowing us to automate the classification scheme in \name.

@@

We now present a meta-theoretic result that certifies the soundness of
classification-based contract enforcement. To help us state the result, we
define an axiomatic model of the our system described informally in
\S~\ref{sec:sysmod}:

### 2014-11-05 evaluation.tex
Edits throughout

\section{Evaluation}

@@

\item \textbf{RUBiS}: An ebay-like auction site~\cite{}. The application allows
users to browse items, bid for items on sale, and pay for items from a wallet
modelled after a bank account.

\item \textbf{Microblog}: A twitter-like microblogging site, modelled after
Twissandra~\cite{}. The application allows adding a new user, adding and
replying to tweets, following, unfollowing and blocking users, and fetching a
user's timeline, userline, followers and following.
\end{itemize}

@@

Figure~\ref{grf:BA-tp-vs-lat} shows the throughtput vs. latency of operations
in bank account example as we increase the number of clients in \cf{1DC}
configuration. Our client workload was generated using YCSB benchmark~\cite{}.
The benchmark uniformly chose from 100,000 keys, where the operation spread was
25\% withdraw, 25\% deposit and 50\% getBalance, which corresponds to the
default 50:50 read:write mix in YCSB. We increased the number of clients from
128 to 1024, and each experiment ran for 180 seconds.

@@

Finally, we study the impact of summarization in
Figure~\ref{grf:summarization}. We utilize 128 clients and a single \name
replica, with all the clients operating on the \emph{same} LWW register to
stress test the summarization mechanism. We utilize 50:50 read:write mix, and
33:33:33 EC:CC:SC mix. The shim layer cache (mem) of operations is summarized
every 64 updates, while the updates in the backing store (disk) are summarized
every 4096 updates. Each point in the graph represents the average latency of
the previous 1000 operations. Each experiment is run for 60s.

>>>>

\section{Evaluation}
\label{sec:results}

@@

\item \textbf{RUBiS}: An ebay-like auction site~\cite{RUBiS}. The application
allows users to browse items, bid for items on sale, and pay for items from a
wallet modelled after a bank account.

\item \textbf{Microblog}: A twitter-like microblogging site, modelled after
Twissandra~\cite{Twissandra}. The application allows adding a new user, adding
and replying to tweets, following, unfollowing and blocking users, and fetching
a user's timeline, userline, followers and following.
\end{itemize}

@@

Figure~\ref{grf:BA-tp-vs-lat} shows the throughtput vs. latency of operations
in bank account example as we increase the number of clients in \cf{1DC}
configuration. Our client workload was generated using YCSB
benchmark~\cite{YCSB}. The benchmark uniformly chose from 100,000 keys, where
the operation spread was 25\% withdraw, 25\% deposit and 50\% getBalance, which
corresponds to the default 50:50 read:write mix in YCSB. We increased the
number of clients from 128 to 1024, and each experiment ran for 180 seconds.

@@

Finally, we study the impact of summarization in
Figure~\ref{grf:summarization}. We utilize 128 clients and a single \name
replica, with all the clients operating on the \emph{same} LWW register to
stress test the summarization mechanism. The shim layer cache (mem) of
operations is summarized every 64 updates, while the updates in the backing
store (disk) are summarized every 4096 updates. Each point in the graph
represents the average latency of the previous 1000 operations. Each experiment
is run for 60s.

### 2014-11-05 implementation.tex
Major addition to implementation
Edits throughout

\section{Implementation}

\name is implemented as a shallow extension of GHC Haskell and runs on top
of Cassandra, an off-the-shelf eventually consistent distributed data (or
backing) store responsible for all data management issues (i.e.,
replication, fault tolerance, availability, and convergence).  Template
Haskell is used implement static contract classification, and proof
obligations are discharged with the help of the Z3~\cite{} SMT
solver. Figure~\ref{fig:impl_mod} describes the overall system architecture.

\begin{figure}
\begin{center}
\includegraphics[width=0.9\columnwidth]{Figures/ImplModel}
\end{center}
\caption{Implementation Model.}
\label{fig:impl_mod}
\end{figure}

>>>>

\section{Implementation}
\label{sec:impl}

\name is implemented as a shallow extension of GHC Haskell and runs on top of
Cassandra, an off-the-shelf eventually consistent distributed data (or backing)
store responsible for all data management issues (i.e., replication, fault
tolerance, availability, and convergence).  Template Haskell is used implement
static contract classification, and proof obligations are discharged with the
help of the Z3~\cite{Z3} SMT solver. Figure~\ref{fig:impl_mod} illustrates the
overall system architecture.

\begin{figure}
\begin{center}
\includegraphics[width=0.7\columnwidth]{Figures/ImplModel}
\end{center}
\caption{Implementation Model.}
\label{fig:impl_mod}
\end{figure}

### 2014-11-05 intro.tex
edits
Edits throughout

Many real-world web services --- such as those built and maintained by
Amazon, Facebook, Google, Twitter, etc. --- replicate application state and
logic across multiple \emph{replicas} within and across data
centers. Replication is intended not only to improve application throughput
and reduce user-perceived latency, but also to tolerate partial failures
without compromising overall service availability. Traditionally programmers
have relied on \emph{strong consistency} guarantees such as
linerarizability~\cite{Herlihy1990} or
serializability~\cite{Serializability} in order to build correct
applications.  While strong consistency is an easily stated property, it
masks the reality underlying large-scale distributed systems with respect to
non-uniform latency (i.e., availability) and network
partitions~\cite{Brewer2000,Gilbert}. Indeed, modern web services, which aim
to provide an "always on" experience, overwhelmingly favour availability and
partition tolerance over strong consistency. To this end, several \emph{weak
  consistency} models such as eventual consistency, causal consistency,
session guarantees, and timeline consistency have been proposed.

Under weak consistency, the developer needs to be aware of concurrent
conflicting updates, and has to pay careful attention to avoid unwanted
inconsistencies (e.g., negative balances in a bank account, or having an
item appear in a shopping cart after it has been
removed~\cite{Dynamo}). Oftentimes, the inconsistency leaks from the
application and is witnessed by the user.  Ultimately, the developer must
decide the consistency level appropriate for a particular operation; this is
understandably an error-prone process requiring intricate knowledge of both
the application as well as the semantics and implementation of the
underlying data store, which typically have only informal
descriptions~\cite{}.  Nonetheless, picking the correct consistency level is
critical not only for correctness but also for scalability of the
application. While choosing a weaker consistency level than required may
introduce program errors and anomalies, choosing a stronger one than
necessary can negatively impact program scalability.

%--KC--
% The choice of correct consistency level is critical to the correctness and
% scalability of the application. Choosing a weaker consistency level than
% required introduces anomalies, while choosing a stronger than necessary level
% affects program scalability.
%\KC{Do we include the LWW-Register number here to add emphasis to the fact that
%choosing the correct consistency level is essential for program scalability}.

@@

The rest of the paper is organized as follows. The next section describes
the system model. Section~\ref{sec:motivation} describes the challenges in
programming with eventually consistent data stores, and introduces
\name\ contracts as a proposed solution to overcome these issues.
Section~\ref{sec:lang} provides more details the contract language.
Section~\ref{sec:opsem} presents an operational semantics.
Section~\ref{sec:txns} introduces transaction contracts and classification.
Section~\ref{sec:impl} describes the implementation and provides details 
about the optimizations needed to make the system practical.
Section~\ref{sec:results} discusses experimental evaluation.
Section~\ref{sec:related} and~\ref{sec:concl} present related work and
conclusions.

>>>>

Many real-world web services --- such as those built and maintained by
Amazon, Facebook, Google, Twitter, etc. --- replicate application state and
logic across multiple \emph{replicas} within and across data
centers. Replication is intended not only to improve application throughput
and reduce user-perceived latency, but also to tolerate partial failures
without compromising overall service availability. Traditionally programmers
have relied on \emph{strong consistency} guarantees such as
linerarizability~\cite{Herlihy1990} or
serializability~\cite{Serializability} in order to build correct
applications.  While strong consistency is an easily stated property, it
masks the reality underlying large-scale distributed systems with respect to
non-uniform latency (i.e., availability) and network
partitions~\cite{Brewer2000,Gilbert2002}. Indeed, modern web services, which aim
to provide an "always on" experience, overwhelmingly favour availability and
partition tolerance over strong consistency. To this end, several \emph{weak
  consistency} models such as eventual consistency, causal consistency,
session guarantees, and timeline consistency have been proposed.

Under weak consistency, the developer needs to be aware of concurrent
conflicting updates, and has to pay careful attention to avoid unwanted
inconsistencies (e.g., negative balances in a bank account, or having an item
appear in a shopping cart after it has been removed~\cite{Dynamo}). Oftentimes,
the inconsistency leaks from the application and is witnessed by the user.
Ultimately, the developer must decide the consistency level appropriate for a
particular operation; this is understandably an error-prone process requiring
intricate knowledge of both the application as well as the semantics and
implementation of the underlying data store, which typically have only informal
descriptions. Nonetheless, picking the correct consistency level is critical
not only for correctness but also for scalability of the application. While
choosing a weaker consistency level than required may introduce program errors
and anomalies, choosing a stronger one than necessary can negatively impact
program scalability.

@@

The rest of the paper is organized as follows. The next section describes the
system model.  We describe the challenges in programming with eventually
consistent data stores, and introduces \name\ contracts as a proposed solution
to overcome these issues in \S~\ref{sec:motivation}. \S~\ref{sec:lang} provides
more details the contract language, and its mapping to the store consistency
levels, along with metatheory for certifying the correctness of the mapping.
\S~\ref{sec:txns} introduces transaction contracts and classification.
\S~\ref{sec:impl} describes the implementation and provides details about the
optimizations needed to make the system practical. \S~\ref{sec:results}
discusses experimental evaluation. \S~\ref{sec:related} and~\ref{sec:concl}
present related work and conclusions.

### 2014-11-05 transactions.tex
Edits throughout

\section{Transaction Contracts}

While contracts on individual operations offer the programmer object-level
declarative reasoning, real-world scenarios often involve operations that
span multiple objects. In order to address this problem, several recent
systems~\cite{COPS,BurckhardtESOP,BailisHAT} have proposed a variety of
transactions, with varying semantics, in order to compose operations on
multiple objects. However, given that classical transaction models such as
serializability~\cite{} and snapshot isolation~\cite{} require inter-replica
coordination, these systems espouse \emph{coordination-free transactions}
that remain available under network partitions, but only provide weaker
isolation guarantees. Coordination-free transactions have intricate
consistency semantics and widely varying runtime overheads. As with
operation-level consistency, the onus is on the programmer to pick the
correct transaction kind; this selection is complicated by operation
consistency requirements.

@@

\noindent \begin{minipage}[t]{0.5\columnwidth}
\begin{codehaskell}
save amt =
  x <- dollar(classify $\psi_1$)
  atomically x dollar do
    b <- withdraw c amt
    when b dollar deposit s amt
\end{codehaskell}
\end{minipage}
\begin{minipage}[t]{0.5\columnwidth}
\begin{codehaskell}
totalBalance =
  x <- dollar(classify $\psi_2$)
  atomically x dollar do
    b1 <- getBalance s
    b2 <- getBalance s
    return b1 + b2
\end{codehaskell}
\end{minipage}

@@

In order to illustrate the utility of transaction contract classification, we
identify three well-understood coordination-free transaction semantics -- Read
Committed (RC)~\cite{}, Monotonic Atomic View (MAV)~\cite{} and Repeatable Read
(RR)~\cite{}, and illustrate the classification strategy. Our technique can
indeed be applied to a different isolation level lattice.

>>>>

\section{Transaction Contracts}
\label{sec:txns}

While contracts on individual operations offer the programmer object-level
declarative reasoning, real-world scenarios often involve operations that span
multiple objects. In order to address this problem, several recent
systems~\cite{Walter,Burckhardt2012,BailisHAT} have proposed a variety of
transactions, with varying semantics, in order to compose operations on
multiple objects. However, given that classical transaction models such as
serializability and snapshot isolation require inter-replica coordination,
these systems espouse \emph{coordination-free transactions} that remain
available under network partitions, but only provide weaker isolation
guarantees. Coordination-free transactions have intricate consistency semantics
and widely varying runtime overheads. As with operation-level consistency, the
onus is on the programmer to pick the correct transaction kind; this selection
is complicated by operation consistency requirements.

@@

\noindent \begin{minipage}[t]{0.53\columnwidth}
\begin{codehaskell}
save amt =
  x <- dollar(classify $\psi_1$)
  atomically x dollar do
    b <- withdraw c amt
    when b dollar deposit s amt
\end{codehaskell}
\end{minipage}
\begin{minipage}[t]{0.47\columnwidth}
\begin{codehaskell}
totalBalance =
  x <- dollar(classify $\psi_2$)
  atomically x dollar do
    b1 <- getBalance s
    b2 <- getBalance s
    return b1 + b2
\end{codehaskell}
\end{minipage}

@@

In order to illustrate the utility of transaction contract classification, we
identify three well-understood coordination-free transaction semantics -- Read
Committed (RC)~\cite{Berenson95}, Monotonic Atomic View (MAV)~\cite{BailisHAT}
and Repeatable Read (RR)~\cite{Berenson95}, and illustrate the classification
strategy. Our technique can indeed be applied to a different isolation level
lattice.
